### 2022-12-30 response.tex

\noindent Dear Editor:

Following an editorial suggestion from PNAS, we are transferring our manuscript ``Intermediate levels of asymptomatic transmission can lead to the highest epidemic fatalities'' for consideration for publication in PNAS Nexus.
Our revised submission builds on multiple years of empirical evidence that asymptomatic transmission of SARS-CoV-2 impacts pandemic controllability and severity. 
Here, we provide and apply a significant refinement of the associated theory, linking asymptomatic transmission at individual scales to population-level epidemic outcomes.

Consistent with PNAS reviewer suggestions, we have revised our manuscript to better highlight the theoretical nature of our work and the associated uncertainties in parametric assumptions.  Notably, we perform sensitivity analyses to show that our central conclusions are robust to parameter variation associated with asymptomatic transmissibility (Supplementary Figure S1--2). We also provide directions for future areas of research that can assess both the impact of asymptomatic transmission and the extent to which control can improve population-level outcomes. Our original conclusions remain: we find that intermediate levels of asymptomaticity lead to the highest levels of epidemic fatalities over a wide range of parameters. Hence, asymptomatic infection and transmission presents a double-edged sword.

As context, foundational analyses of asymptomatic infections have primarily focused on the impact of asymptomatic infections on disease controllability (e.g., Fraser et al., 2004) – neglecting the effects of asymptomatic spread on population-level outcomes, such as overall fatalities. Here, we use a series of epidemic models to assess the role of asymptomatic spread on epidemic fatalities in the presence of transmission-reducing effects among symptomatic individuals (e.g., via behavioural change or symptom-based interventions). We then extend our framework to assess the ways in which immunity profiles can lead to potentially paradoxical outcomes: with less virulent strains spreading precisely because asymptomatic individuals can transmit while unaware they are infected.   Moreover, our findings show that emergence of milder variants at the individual scale can nonetheless cause similar or even higher levels of fatalities at population scales – mirroring many features of the Omicron variant. Overall, we believe that the revised work is timely, novel, and fits the PNAS Nexus mission of welcoming rigorous research with broad, interdisciplinary appeal.
